Proof nets for first-order additive linear logic

### 2017-12-19 ALL1.tex
initial upload - 10 pages

\newcommand\subs[1]{\textsc{sub}(#1)}
\newcommand\dom[1]{\textsc{dom}(#1)}

@@

\newcommand\sub{\leq}
\newcommand\seq[3][]{{\vdash_{#1}}#2,#3}
\newcommand\fv{\textsc{fv}}

@@

\newcommand\link[3][\sigma]{(#2,#3)_{#1}}
\newcommand\minus{\mathop{/\mathchoice{\kern-3pt}{\kern-3pt}{\kern-2.5pt}{\kern-2pt}/}}

@@

\[
\begin{tikzpicture}[net]
\formula[y=1]{\exists x.\forall a.~p(a,x)}
\formula[y=0]{\exists y.~{\dual p}(y,s)\*{\dual p}(y,t)}
\Vlink[red,label={$\quad\scriptstyle{\upsilon\!\minus\!y}$}]{8,1}
\end{tikzpicture}
\]

\[
\begin{tikzpicture}[net]
\formula[y=1]{\exists x.\forall a.~p(a,x)}
\formula[y=0]{\exists y.~{\dual p}(y,s)\*{\dual p}(y,t)}
\Vlink[red,label={$~\scriptstyle{\upsilon\!\minus\!y}$}]{4,1}
\end{tikzpicture}
\]

@@

\begin{figure}
\[
	\infer{\seq P{\dual P}}{}
\qquad
	\infer{\seq A{B_0\+B_1}}{\seq A{B_i}}
\qquad
	\infer{\seq A{B\*C}}{\seq AB && \seq AC}
\qquad
	\infer{\seq A{\exists x.B}}{\seq A{B[t/x]}}
\qquad
	\infer[\!\!\scriptstyle{a\,\notin\,\textsc{fv}(A)}]{\seq A{\forall a.B}}{\seq AB}
\]
\caption{A sequent calculus for $\all1$}
\label{fig:sequent calculus}
\end{figure}

\begin{figure}
\noindent\hrule
\[
	\infer[\!\!\scriptstyle{\qrr1}]{\seq P{\dual P}}{}
\quad
	\infer[\!\!\scriptstyle{\qrr+i}]{\seq A{B_0\+B_1}}{\seq A{B_i}}
\quad
	\infer[\!\!\scriptstyle{\qrr*}]{\seq A{B\*C}}{\seq AB & \seq AC}
\quad
	\infer[\!\!\scriptstyle{\qrr?t}]{\seq A{\exists x.B}}{\seq A{B[t/x]}}
\quad
	\infer[\!\!\scriptstyle{\qrr!~(a\,\notin\,\textsc{fv}(A))}]{\seq A{\forall a.B}}{\seq AB}
\]
\caption{A sequent calculus for $\all1$}
\label{fig:sequent calculus}
\end{figure}
%
%\begin{figure}
%\[
%\raisebox{.5ex}{$\qrr1$}\quad
%	\infer{\seq P{\dual P}}{}
%\quad\raisebox{.5ex}{$\qrr+i$}\quad
%	\infer{\seq A{B_0\+B_1}}{\seq A{B_i}}
%\quad\raisebox{.5ex}{$\qrr*$}\quad
%	\infer{\seq A{B\*C}}{\seq AB && \seq AC}
%\quad\raisebox{.5ex}{$\qrr?$}\quad
%	\infer{\seq A{\exists x.B}}{\seq A{B[t/x]}}
%\quad\raisebox{.5ex}{$\qrr!$}\quad
%	\infer[\!\!\scriptstyle{a\,\notin\,\textsc{fv}(A)}]{\seq A{\forall a.B}}{\seq AB}
%\]
%\caption{A sequent calculus for $\all1$}
%\label{fig:sequent calculus}
%\end{figure}

@@

First-order terms and the formulas of first-order $\all$ are generated by the following grammars.
%
\setMidspace{5pt}
\[
\begin{array}{@{}l@{}l}
	t &\Coloneqq a \Mid x \Mid f(t_1,\dots,t_n)
\\[10pt]
	P &\Coloneqq p(t_1,\dots,t_n) \Mid \dual p(t_1,\dots,t_n)
\\[10pt]
	A &\Coloneqq P \Mid A\+A \Mid A\*A \Mid \exists x.A \Mid \forall a.A
\end{array}
\]
%
Negation $(\dual{\,\cdot\,})$ is applied to predicate symbols, $\dual p$, as a matter of convenience. The \defn{dual} $\dual A$ of an arbitrary formula $A$ is given by DeMorgan. We use the following notational conventions.
%
\[
\begin{tabular}{@{}lll@{}}
	$a,b,c$ & $\in \varA$ 		& universally quantified variables (and free variables)\\
	$x,y,z$ & $\in \varE$		& existentially quantified variables \\
	$f,g,h$ & $\in \Sigma_f$	& $n$-ary $(n\geq 0)$ function symbols from a fixed alphabet $\Sigma_f$\\
	$p,q,r$ & $\in \Sigma_p$	& $n$-ary $(n\geq 0)$ predicate symbols from a fixed alphabet $\Sigma_p$ \\
	$s,t,u$ & $\in \terms$ 		& first-order terms over $\varA$, $\varE$, $\Sigma_f$, and $\Sigma_p$ \\
	$P,Q,R$ & $\in \atom$		& atomic propositions \\
	$A,B,C$ & $\in \form$		& \all1 formulas \\
\end{tabular}
\]
%
A \defn{sequent} $\seq AB$ is a pair of formulas $A$ and $B$. A sequent calculus for \all1 is given in Figure~\ref{fig:sequent calculus}, where each rule has a symmetric counterpart that applies to the first formula in the sequent. We write $\prf\pi AB$ for a proof $\pi$ with conclusion sequent $\seq AB$. Anticipating the requirements of proof nets, we will assume the following convention for \all1 proofs.

\begin{description}
	\item
[\defn{Barendregt's convention}] All quantifiers in a sequent $\seq AB$ have a distinct binding variable, and no bound variable shares a name with a free one.
%	\item
%% ALTERNATIVE: "STATIC WITNESSING"
%[\defn{Universal witnessing}] All existential witnesses in a proof are \defn{universal terms} $\termsA\subset\terms$, generated by the following grammar.
%\[
%	\termsA: \qquad t \Coloneqq a \Mid f(t_1,\dots,t_n)
%\]
\end{description}

@@

A \defn{witness map} $\sigma\colon \varE \rightharpoonup \terms$ is a substitution map which assigns terms to existential variables, given as a (finite) partial function. The \defn{empty} witness map $\varnothing$ is the everywhere undefined partial function, and the map $\sigma[t/x]$ substitutes $t$ for $x$ and is as $\sigma$ everywhere else. We express by $\sigma(x)=\bot$ that $\sigma$ is undefined for $x$. The map $\sigma\minus x$ is undefined on $x$ and otherwise as $\sigma$.

We will write $A\sigma$ for the application of the witness map $\sigma\colon\varE\rightharpoonup\terms$ to the formula $A$ as a substitution map. It is applied to a sequent proof, $\pi\sigma$, by applying it to each formula in the proof $\pi$. The \defn{composition} of two maps is written $\sigma\tau$, where $A(\sigma\tau)=(A\sigma)\tau$. We define the following further operations on witness maps.

\[
\begin{tabular}{@{}lp{.75\textwidth}@{}}
	$\sigma\gen\tau$
&
	A witness map is \defn{more general} than another, 
	$\sigma\gen\tau$, if there is a map $\rho$ such that 
	$\sigma\rho=\tau$.
\\[5pt]	
	$\sigma\coh\tau$
&
	Two witness maps are \defn{coherent}, $\sigma\coh\tau$,
	if there is a map $\rho$ such that $\sigma\rho=\tau\rho$.
\\[5pt]
	$\sigma\join\tau$
&
	The \defn{join} of two witness maps $\sigma\join\tau$ is defined
	if they are coherent, and is the least map $\rho$ such that
	$\sigma\gen\rho$ and $\tau\gen\rho$.
\end{tabular}%
\]

A \defn{witness linking} $\lambda_\Sigma$ is a linking $\lambda$ with a \defn{witness labelling}  $\Sigma\colon\lambda\to \varE\rightharpoonup\terms$ that assigns each link $(C,D)$ a witness map. 
We may use and define $\lambda_\Sigma$ as a set of \defn{witness links} $\link CD$, which indicates the link $(C,D)$ in $\lambda$ where $\Sigma(C,D)=\sigma$.

\begin{definition}
A \defn{witness pre-net} $\net{\lambda_\Sigma}AB$ is a pre-net with a witness linking $\lambda_\Sigma$.
\end{definition}

We assume the existential variables occurring in the range of $\Sigma$ to be fresh, i.e.\ distinct from any existentially quantified variable in $\seq AB$. We abbreviate by $a\in \sigma$ the statement that a (universal) variable $a$ occurs free in the range of $\sigma$, i.e.\ that $a\in\fv(\sigma(x))$ for some $x$.

@@

A witness link $\link PQ$ or link $(P,Q)$ on two atomic formulas $P$ and $Q$ is an \defn{axiom} link if $P\sigma=\dual Q\sigma$, or respectively, if some $\sigma$ exists such that $P\sigma=\dual Q\sigma$. An \defn{axiom} (witness) linking is one consisting only of axiom links. Observe that in the de-sequentialization $[\pi]$ the witness linking $\deseq[\varnothing]\pi AB$ is an axiom linking.

To an axiom link $(P,Q)$ we assign an \defn{initial witness map} $\sigma=\init PQ$ as the least witness map in $(\gen)$ such that $P\sigma=\dual Q\sigma$ --- in other words, $\init PQ$ is the most general unifier ($\textsc{mgu}$) of $P$ and $\dual Q$. An axiom linking $\lambda$ with the initial witness map assigned to each link will be written $\lambda_\star$:
\[
	\lambda_\star = \{~\link PQ~\mid~(P,Q)\in\lambda~,~\sigma=\init PQ~\}~.
\]

For two witness labellings $\Sigma$ and $\Xi$ on the same linking $\lambda$, one is \defn{more general} than the other, $\Sigma\gen\Xi$, if $\Sigma(C,D)\gen\Xi(C,D)$ for every link $(C,D)\in\lambda$.  $\lambda_\Sigma\gen\lambda_\Xi$,

Observe that for an axiom linking $\lambda$, the initial linking $\lambda_\star$ is the most general among the witness linkings $\lambda_\Sigma$ for any $\Sigma$.

@@

We will define two coalescence relations, one \defn{strict} $(\scoal)$ and one \defn{unifying} $(\ucoal)$. In the former, to join two slices the witness assignment must be the same; in the latter, witness assignments must be compatible, so that a common, more general witness map exists.

For sequentialization, the links in a pre-net will be labelled with a sequent proof. An axiom link will carry an axiom, and each coalescence step introduces one proof rule. A \defn{proof linking} $\lambda_\Sigma^\Pi$ is a witness linking $\lambda_\Sigma$ with a \defn{proof labelling} $\Pi\colon \lambda\to\proofs$ assigning each link a sequent proof. We will use and define $\lambda_\Sigma^\Pi$ as a set of \defn{proof links} $\link CD^\pi$, where we assume that $\prf\pi{C\sigma}{D\sigma}$.

Let a \defn{labelled link} $\link AB^\pi$ be a link carrying a sequent proof $\prf\pi{A\sigma}{B\sigma}$, and a \defn{labelled linking} or \defn{labelled pre-net} one where every link is labelled. The \defn{initial labelling} $\lambda_\Sigma^\star$ of an axiom linking $\lambda_\Sigma$ is:
\[
	\lambda_\Sigma^\star = \{~\link PQ^\pi~\mid~\link PQ\in\lambda_\Sigma~,~\pi=~\infer{\seq{P\sigma}{Q\sigma}}{}~\}~.
\]
For correctness we may coalesce a pre-net without constructing a proof. 
%
To recap, we have accumulated the following further notational conventions.
%
\[
\begin{tabular}{@{}lcll@{}}
	$\kappa,\lambda$ 	& $\subset$ & $\form\times\form$ 	& linkings (sets of pairs of formulas)\\
	$\rho,\sigma,\tau$	& $\colon$  & $\varE\rightharpoonup\terms$ & witness maps\\
	$\Sigma,\Xi$		& $\colon$  & $(\form\times\form)\rightharpoonup\varE\rightharpoonup\terms$		& witness labellings on a linking \\
	$\pi,\phi,\psi$		& $\in$ & $\proofs$ & \all1 sequent proofs\\
	$\Pi,\Phi,\Psi$		& $\colon$  & $(\form\times\form)\rightharpoonup\proofs$		& proof labellings on a linking
\end{tabular}
\]

\begin{definition}
 \defn{Strict sequentialization} ($\scoal$) is the rewrite relation on labelled pre-nets generated by the following rules, which apply symmetrically to the first formula in the sequent. \defn{Strict coalescence} is the same relation ignoring proof labels, on witness pre-nets.
%
\renewcommand{\+}{{+}}
\renewcommand{\*}{{\times}}
\begin{equation}
  	\net{\{\link A{B_i}^\pi\}}A{B_0\+B_1} 
 ~\scoal~
	\net{\{\link A{B_0\+B_1} ^\psi\}}A{B_0\+B_1}
\tag{$\srr+i$}
\end{equation}

@@

\begin{equation}
	\net{\{\link AB^\pi\}}A{\exists x.B}
 ~\scoal~ 
	\net{\{\link[\sigma\!\minus\!x]A{\exists x.B}^\psi\}}A{\exists x.B}
\tag{$\srr?$}
\end{equation}

@@

\noindent
A pre-net $\net{\lambda_\Sigma}AB$ \defn{strict-coalesces} if it reduces in $(\scoal)$ to a single link $\net{\{\link[\varnothing]AB\}}AB$.
\end{definition}

To be explicit, to generate coalescence as a rewrite relation means the following: a pre-net coalesces to another, $\net{\lambda_\Sigma} AB\scoal \net{{\lambda'_{\Sigma'}}}AB$ if and only if % there is a rewrite rule $\net{\kappa_\Xi} CD\scoal\net{\kappa'_{\Xi'}} CD$ such that $C\leq A$, $D\leq B$, ${\kappa_\Xi}\subseteq{\lambda_\Sigma}$, $\kappa'_{\Xi'}\subseteq{\lambda'_{\Sigma'}}$, and ${\lambda_\Sigma}-{\kappa_\Xi}={\lambda'_{\Sigma'}}-\kappa'_{\Xi'}$.
there is a rewrite rule
\[
	\net{\kappa_\Xi} CD\scoal\net{\kappa'_{\Xi'}} CD
\] 
such that 
\[
	C\leq A \qquad
	D\leq B \qquad
	\kappa_\Xi\subseteq\lambda_\Sigma \qquad
	\kappa'_{\Xi'}\subseteq\lambda'_{\Sigma'} \quad
	\text{and} \quad
	\lambda_\Sigma-\kappa_\Xi=\lambda'_{\Sigma'}-\kappa'_{\Xi'}~.
\]

>>>>

\newcommand\ex[2]{\textsc{ex}_{#1}(#2)}

\newcommand\subs[1]{\textsc{sub}(#1)}
\newcommand\poss[1]{\textsc{pos}(#1)}
\newcommand\dom[1]{\textsc{dom}(#1)}

@@

\newcommand\sub{\leq}
\newcommand\dep{\preccurlyeq}
%\newcommand\dep{\stackrel\star{\smash\preccurlyeq\rule{0pt}{1ex}}}

\newcommand\seq[3][]{{\vdash_{#1}}#2,#3}
\newcommand\fv{\textsc{fv}}

@@

\newcommand\link[3][\sigma]{(#2,#3)_{#1}}
\newcommand\minus{\mathop{\!/\mathchoice{\kern-3pt}{\kern-3pt}{\kern-2.5pt}{\kern-2pt}/\!}}

@@

\[
\begin{tikzpicture}[net]
\formula[y=1]{\exists x.\forall a.~p(a,x)}
\formula[y=0]{\exists y.~{\dual p}(y,s)\*{\dual p}(y,t)}
\Vlink[red,label={$\quad\scriptstyle{\upsilon\minus y}$}]{8,1}
\end{tikzpicture}
\]

\[
\begin{tikzpicture}[net]
\formula[y=1]{\exists x.\forall a.~p(a,x)}
\formula[y=0]{\exists y.~{\dual p}(y,s)\*{\dual p}(y,t)}
\Vlink[red,label={$~\scriptstyle{\upsilon\minus y}$}]{4,1}
\end{tikzpicture}
\]

@@

%\begin{figure}
%\[
%	\infer{\seq P{\dual P}}{}
%\qquad
%	\infer{\seq A{B_0\+B_1}}{\seq A{B_i}}
%\qquad
%	\infer{\seq A{B\*C}}{\seq AB && \seq AC}
%\qquad
%	\infer{\seq A{\exists x.B}}{\seq A{B[t/x]}}
%\qquad
%	\infer[\!\!\scriptstyle{a\,\notin\,\textsc{fv}(A)}]{\seq A{\forall a.B}}{\seq AB}
%\]
%\caption{A sequent calculus for $\all1$}
%\label{fig:sequent calculus}
%\end{figure}

\begin{figure}
\[
	\infer[\!\!\scriptstyle{\qrr1}]{\seq P{\dual P}}{}
\quad
	\infer[\!\!\scriptstyle{\qrr+i}]{\seq A{B_0\+B_1}}{\seq A{B_i}}
\quad
	\infer[\!\!\scriptstyle{\qrr*}]{\seq A{B\*C}}{\seq AB & \seq AC}
\quad
	\infer[\!\!\scriptstyle{\qrr?t}]{\seq A{\exists x.B}}{\seq A{B[t/x]}}
\quad
	\infer[\!\!\scriptstyle{\qrr!~(a\,\notin\,\textsc{fv}(A))}]{\seq A{\forall a.B}}{\seq AB}
\]
\caption{A sequent calculus for $\all1$}
\label{fig:sequent calculus}
\end{figure}
%
%\begin{figure}
%\[
%\raisebox{.5ex}{$\qrr1$}\quad
%	\infer{\seq P{\dual P}}{}
%\quad\raisebox{.5ex}{$\qrr+i$}\quad
%	\infer{\seq A{B_0\+B_1}}{\seq A{B_i}}
%\quad\raisebox{.5ex}{$\qrr*$}\quad
%	\infer{\seq A{B\*C}}{\seq AB && \seq AC}
%\quad\raisebox{.5ex}{$\qrr?$}\quad
%	\infer{\seq A{\exists x.B}}{\seq A{B[t/x]}}
%\quad\raisebox{.5ex}{$\qrr!$}\quad
%	\infer[\!\!\scriptstyle{a\,\notin\,\textsc{fv}(A)}]{\seq A{\forall a.B}}{\seq AB}
%\]
%\caption{A sequent calculus for $\all1$}
%\label{fig:sequent calculus}
%\end{figure}

@@

First-order terms and the formulas of first-order $\all$ are generated by the following grammars.
%
\setMidspace{5pt}
\[
\begin{array}{@{}l@{}l}
	t &\Coloneqq a \Mid x \Mid f(t_1,\dots,t_n)
\\[10pt]
	P &\Coloneqq p(t_1,\dots,t_n) \Mid \dual p(t_1,\dots,t_n)
\\[10pt]
	A &\Coloneqq P \Mid A\+A \Mid A\*A \Mid \exists x.A \Mid \forall a.A
\end{array}
\]
%
Negation $(\dual{\,\cdot\,})$ is applied to predicate symbols, $\dual p$, as a matter of convenience. The \defn{dual} $\dual A$ of an arbitrary formula $A$ is given by DeMorgan. We use the following notational conventions.
%
\[
\begin{tabular}{@{}lll@{}}
	$a,b,c$ & $\in \varA$ 		& universally quantified variables (and free variables)\\
	$x,y,z$ & $\in \varE$		& existentially quantified variables (and term-language variables)\\
	$f,g,h$ & $\in \Sigma_f$	& $n$-ary $(n\geq 0)$ function symbols from a fixed alphabet $\Sigma_f$\\
	$p,q,r$ & $\in \Sigma_p$	& $n$-ary $(n\geq 0)$ predicate symbols from a fixed alphabet $\Sigma_p$ \\
	$s,t,u$ & $\in \terms$ 		& first-order terms over $\varA$, $\varE$, $\Sigma_f$, and $\Sigma_p$ \\
	$P,Q,R$ & $\in \atom$		& atomic propositions \\
	$A,B,C$ & $\in \form$		& \all1 formulas \\
\end{tabular}
\]
%
A \defn{sequent} $\seq AB$ is a pair of formulas $A$ and $B$. A sequent calculus for \all1 is given in Figure~\ref{fig:sequent calculus}, where each rule has a symmetric counterpart that applies to the first formula in the sequent. We write $\prf\pi AB$ for a proof $\pi$ with conclusion sequent $\seq AB$.

@@

A \defn{witness map} $\sigma\colon \varE \rightharpoonup \terms$ is a substitution map which assigns terms to existential variables, given as a (finite) partial function. Its \defn{domain} $\dom(\sigma)$ are those variables for which it is defined. The \defn{empty} witness map $\varnothing$ is the everywhere undefined partial function, and the map $\sigma[t/x]$ substitutes $t$ for $x$ and is as $\sigma$ everywhere else. We express by $\sigma(x)=\bot$ that $\sigma$ is undefined for $x$. The map $\sigma\minus x$ is undefined on $x$ and otherwise as $\sigma$.

We will write $A\sigma$ for the application of the witness map $\sigma\colon\varE\rightharpoonup\terms$ to the formula $A$ as a substitution map. It is applied to a sequent proof, $\pi\sigma$, by applying it to each formula in the proof $\pi$. The \defn{composition} of two maps is written $\sigma\tau$, where $A(\sigma\tau)=(A\sigma)\tau$.

A \defn{witness linking} $\lambda_\Sigma$ is a linking $\lambda$ with a \defn{witness labelling} $\Sigma\colon\lambda\to \varE\rightharpoonup\terms$ that assigns each link $(C,D)$ a witness map. We may use and define $\lambda_\Sigma$ as a set of \defn{witness links} $\link CD$, where $(C,D)$ in $\lambda$ and $\Sigma(C,D)=\sigma$. When link is not a witness link may be emphasized by calling it \defn{clean}.

\begin{definition}
A \defn{witness pre-net} $\net{\lambda_\Sigma}AB$ is a sequent $\seq AB$ with a witness linking $\lambda_\Sigma$.
%
%\begin{description}
%	\item[Exact coverage]
%The domain of $\sigma$ is exactly the variables existentially quantified in $\seq AB$ that will be bound in $C$ and $D$, if they occur:
%\[
%	\dom\sigma = \{\,x \mid C < \exists x.X \leq A\,\}
%			\cup \{\,y \mid D < \exists y.Y \leq B\,\}~.
%\]
%
%	\item[Eigenvariables not free]
%A universal variable $a\in\varA$ quantified in $C$ or $D$ may not occur free in the range of $\sigma$, $a\notin\sigma$.
%
%	\item[Freshness]
%Existential variables in the range of $\sigma$ are fresh, i.e.\ distinct from any existentially quantified variable in $\seq AB$.
%\end{description}
\end{definition}

We will assume the following variable naming conventions for \all1 pre-nets.

\begin{description}
	\item
[\defn{Barendregt's convention}] All quantifiers in a sequent $\seq AB$ have a distinct binding variable, and no bound variable shares a name with a free one.
%	\item
%% ALTERNATIVE: "STATIC WITNESSING"
%[\defn{Universal witnessing}] All existential witnesses in a proof are \defn{universal terms} $\termsA\subset\terms$, generated by the following grammar.
%\[
%	\termsA: \qquad t \Coloneqq a \Mid f(t_1,\dots,t_n)
%\]

	\item[Eigenvariables not free]
For a link $\link CD$, a universal variable $a\in\varA$ quantified in $C$ or $D$ may not occur free in the range of $\sigma$, $a\notin\sigma$.

	\item[Freshness]
For a link $\link CD$, existential variables in the range of $\sigma$ are fresh, i.e.\ distinct from any existentially quantified variable in $\seq AB$.
\end{description}

%We assume the existential variables occurring in the range of $\Sigma$ to be fresh, i.e.\ distinct from any existentially quantified variable in $\seq AB$.  
We abbreviate by $a\in \sigma$ the statement that a variable $a$ occurs free in the range of $\sigma$, i.e.\ that $a\in\fv(\sigma(x))$ for some $x$.

@@

A witness link $\link PQ$ on two atomic formulas $P$ and $Q$ is an \defn{axiom} link if $P\sigma=\dual Q\sigma$; a witness \defn{axiom} linking is one where every link is an axiom link. Observe that the de-sequentialization $[\pi]$ is a witness pre-net with axiom linking $\deseq[\varnothing]\pi AB$.

@@

We will define two coalescence relations, one \defn{strict} $(\scoal)$, for witness nets, and one \defn{unifying} $(\ucoal)$, for unification nets, in a later section. In the former, to join two slices the witness assignment must be the same; in the latter, witness assignments must be compatible, so that a common, more general witness map exists.

For sequentialization, the links in a pre-net will be labelled with a sequent proof. An axiom link will carry an axiom, and each coalescence step introduces one proof rule. Formalizing this, a \defn{proof linking} $\lambda_\Sigma^\Pi$ is a witness linking $\lambda_\Sigma$ with a \defn{proof labelling} $\Pi\colon \lambda\to\proofs$ assigning each link a sequent proof. We will use and define $\lambda_\Sigma^\Pi$ as a set of \defn{proof links} $\link CD^\pi$, where we require that $\prf\pi{C\sigma}{D\sigma}$, i.e.\ that $\pi$ proves the conclusion $\seq{C\sigma}{D\sigma}$. A \defn{labelled pre-net} $\net{\lambda_\Sigma^\Pi}AB$ is a witness pre-net $\net{\lambda_\Sigma}AB$ with a proof labelling $\Pi$ on $\lambda_\Sigma$. The \defn{initial proof labelling} $\lambda_\Sigma^\star$ of an axiom linking $\lambda_\Sigma$ is:
\[
	\lambda_\Sigma^\star = \{~\link PQ^\pi~\mid~\link PQ\in\lambda_\Sigma~,~\pi=~\infer{\seq{P\sigma}{Q\sigma}}{}~\}~.
\]
For correctness we may coalesce a pre-net directly, without constructing a proof.

To recap, we have accumulated the following further notational conventions.
%
\[
\begin{tabular}{@{}lcll@{}}
	$\kappa,\lambda$ 	& $\subset$ & $\form\times\form$ 	& linkings (sets of pairs of formulas)\\
	$\rho,\sigma,\tau$	& $\colon$  & $\varE\rightharpoonup\terms$ & witness maps\\
	$\Sigma,\Xi$		& $\colon$  & $(\form\times\form)\rightharpoonup\varE\rightharpoonup\terms$		& witness labellings on a linking \\
	$\pi,\phi,\psi$		& $\in$ & $\proofs$ & \all1 sequent proofs\\
	$\Pi,\Phi,\Psi$		& $\colon$  & $(\form\times\form)\rightharpoonup\proofs$		& proof labellings on a linking
\end{tabular}
\]

\begin{definition}
 \defn{Strict sequentialization} ($\scoal$) is the rewrite relation on witness pre-nets generated by the following rules, which apply symmetrically to the first formula in the sequent. \defn{Strict coalescence} is the same relation ignoring proof labels, on witness pre-nets.
%
\renewcommand{\+}{{+}}
\renewcommand{\*}{{\times}}
\begin{equation}
  	\net{\{\link A{B_i}^\pi\}}A{B_0\+B_1} 
 ~\scoal~
	\net{\{\link A{B_0\+B_1} ^\psi\}}A{B_0\+B_1}
\tag{$\srr+i$}
\end{equation}

@@

\begin{equation}
	\net{\{\link AB^\pi\}}A{\exists x.B}
 ~\scoal~ 
	\net{\{\link[\sigma\minus x]A{\exists x.B}^\psi\}}A{\exists x.B}
\tag{$\srr?$}
\end{equation}

@@

\noindent
A witness pre-net $\net{\lambda_\Sigma}AB$ \defn{strict-coalesces} if it reduces to a single link $\net{\{\link[\varnothing]AB\}}AB$.\end{definition}

To be explicit, to generate coalescence as a rewrite relation means the following: a pre-net coalesces to another, $\net{\lambda_\Sigma} AB\scoal \net{{\lambda'_{\Sigma'}}}AB$ if and only if there is a rewrite rule
\[
	\net{\kappa_\Xi} CD\scoal\net{\kappa'_{\Xi'}} CD
\] 
such that 
\[
	C\leq A \qquad
	D\leq B \qquad
	\kappa_\Xi\subseteq\lambda_\Sigma \qquad
	\kappa'_{\Xi'}\subseteq\lambda'_{\Sigma'} \quad
	\text{and} \quad
	\lambda_\Sigma-\kappa_\Xi=\lambda'_{\Sigma'}-\kappa'_{\Xi'}~.
\]

Two useful properties of witness nets are preserved over a coalescence sequence, captured in the lemma below. The first property is formulated the way it is to account for vacuous quantifiers, for which no actual variable occurs in $C$ or $D$, as well as formulas such as $\exists x.C\+C'$ where $x$ occurs in $C'$ but not $C$, where we still would like an explicit witness for $x$ on a link $\link CD$.

\begin{lemma}
Every pre-net on a coalescence path $\net{\lambda_\Sigma}AB \scoal^* \net{\{\link[\varnothing]AB^\pi\}}AB$ from a witness net satisfies the \defn{exact coverage} property: for any link $\link CD$, the domain of $\sigma$ is exactly the variables existentially quantified in $\seq AB$ thad bind in $C$ and $D$,
\[
	\dom\sigma = \{\,x \mid C < \exists x.X \leq A\,\}
			\cup \{\,y \mid D < \exists y.Y \leq B\,\}~.
\]
\end{lemma}

%
%\begin{lemma}
%Every pre-net on a coalescence path $\net{\lambda_\Sigma}AB \scoal^* \net{\{\link[\varnothing]AB^\pi\}}AB$ from a witness net satisfies the following properties. 
%\begin{description}
%	\item[Eigenvariables not free]
%For any link $\link CD$, a universal variable $a\in\varA$ quantified in $C$ or $D$ does not occur free in the range of $\sigma$, $a\notin\sigma$.
%
%	\item[Exact coverage]
%For any link $\link CD$, the domain of $\sigma$ is exactly the variables existentially quantified in $\seq AB$ that will be bound in $C$ and $D$, if they occur:
%\[
%	\dom\sigma = \{\,x \mid C < \exists x.X \leq A\,\}
%			\cup \{\,y \mid D < \exists y.Y \leq B\,\}~.
%\]
%\end{description}
%\end{lemma}

\begin{proof}
By induction on the coalescence path, going forward from $\net{\lambda_\Sigma}AB$ for the first property, and working backward from $\net{\{\link[\varnothing]AB^\pi\}}AB$ for the second.
\end{proof}

\begin{theorem}
For any \all1 proof $\pi$, the de-sequentialization $[\pi]$ is a witness net that sequentializes to $\pi$. 
\end{theorem}

\begin{proof}
It follows by induction on $\prf\pi AB$ that if $\lambda_\Sigma=[\pi]_{A',B'}^\sigma$ where $A'\sigma=A$ and $B'\sigma=B$, then $\net{\lambda^\star_\Sigma}AB$ reduces in $(\scoal)$ to $\net{\{\link{A'}{B'}^\pi\}}AB$. The statement is the specific case $\sigma=\varnothing$.
\end{proof}

%A formula $\exists x.A$ has two natural immediate descendants: one \emph{syntactic}, its subformula $A$, and one \emph{semantic}, the premise $A[t/x]$ for its introduction rule, for any given $t$. We will define the latter as a tree-order generated from a formula $A$: the \defn{positions} of $A$ are its subformulas under substitution,
%\[
%	\poss A~=~\{~B'\sigma~\mid~A=A'\sigma~,~B'\sub A'~,~\sigma\in\varE\rightharpoonup\terms~\}
%\]

A \defn{slice} is the fraction of a proof that depends on a given choice of one branch (or projection) on each product formula $A\*B$. Important to additive proof theory is that many operations can be performed on a per-slice basis, such as normalization, or proof net correctness. We will here use slices for the latter purpose.

As in the propositional case, we define a slice of a sequent $\seq AB$ as a set of potential links, of which exactly one must be realized in a proof net $\net{\lambda_\Sigma}AB$.

%In the propositional case, slices are readily defined as sets of subformulas. Here, in the first-order case, we will need to

%The immediate problem is that the number of positions immediately below $\exists x.B$ is infinite, as there is one immediate descendant $B[t/x]$ for every term $t$. We reduce it to finite width by restricting a slice to just the terms assigned to $x$ by the witness labelling $\Sigma$ in a pre-net $\net{\lambda_\Sigma}AB$.

For a witness linking $\lambda_\Sigma$ and an existential variable $x$, write $\Sigma(x)\subseteq\terms$ for the \defn{witness set} of $x$, which collects the terms assigned to $x$ by the witness maps in $\Sigma$ for every link in $\lambda$:
\[
	\Sigma(x) = \{~\sigma(x)~\mid~\link CD\in\lambda_\Sigma~\}~.
\]
%
Given a witness linking $\lambda_\Sigma$, a \defn{slice} $S$ of a formula $A$ and a witness map $\sigma$ is a set of pairs $(A',\sigma)$ where $A'$ is a subformula of $A$, given by: $S=\{(A,\sigma)\}\cup S'$ where $S'$ is as follows.
%
\begin{itemize}
	\item
\makebox[70pt][l]{If $A = P$}			then $S'=\varnothing$.
	\item
\makebox[70pt][l]{If $A = B\+C$} 		then $S'=S_B\uplus S_C$ where $S_B$ is a slice of $B$ and $\sigma$, and $S_C$ a slice of $C$ and $\sigma$.
	\item
\makebox[70pt][l]{If $A = B\*C$}		then $S'$ is a slice of $B$ and $\sigma$ or a slice of $C$ and $\sigma$.
	\item
\makebox[70pt][l]{If $A = \exists x.B$}	then $S'=\uplus_{t\,\in\,\Sigma(x)}S_t$ where each $S_t$ is a slice of $B$ and $\sigma[t/x]$.
	\item
\makebox[70pt][l]{If $A = \forall a.B$}	then $S'$ is a slice of $B$ and $\sigma$.
\end{itemize}
%
\noindent
A \defn{slice} of a sequent $\seq AB$ is a set of links
\[
	\{~\link[\sigma\cup\tau]CD~\mid~(C,\sigma)\in S_A~,~(D,\tau)\in S_B~\}
\]
where $S_A$ is a slice of $A$ and $\varnothing$, and $S_B$ a slice of $B$ and $\varnothing$.  A \defn{slice} of a witness pre-net $\net{\lambda_\Sigma} AB$ is the intersection $\lambda_\Sigma\cap S$ of $\lambda_\Sigma$ with a slice $S$ of $\seq AB$.

Within a slice, we need to ensure that the order in which quantifiers are instantiated is sound. That is, we need to have an analogue of the \emph{eigenvariable condition} on the $\qrr!$-rule of sequent calculus, that the quantified variable $a$ of the primary formula $\forall a.A$ does not occur free in the context. We will define a \defn{dependency} condition to this effect.

Since we will require that each slice contains a single link, for simplicity we will define the dependency condition on individual links rather than on slices. It is expected that for a more involved logic that includes multiplicative connectives, a dependency condition such as the below one needs to be applied to a slice instead.

For a subformula $C$ of a formula $A$, let the \defn{existential covering} $\ex AC$ of $C$ be the set of variables that are existentially quantified in $A$ dominating $C$:
\[
	\ex AC = \{\,x \mid C < \exists x.X \sub A\,\}
\]
In a pre-net $\net{\lambda_\Sigma}AB$, let the \defn{column} of a link $\link CD$ be the set of pairs 
\[
	\{~(X,\sigma|\ex AX)~\mid~C\sub X\sub A~\}\cup\{~(Y,\sigma|\ex BY)~\mid~D\sub Y\sub B~\}~,
\]
with an order $(\dep)$ given by: $(X,\rho)\dep(Y,\tau)$ if $X\leq Y$, or $Y$ occurs as $\forall a.Y$ and $a\in\rho$.

\begin{definition}
A witness pre-net is \defn{correct} if it is
\begin{enumerate}
\item \defn{slice-correct}: every slice is a singleton, and
\item \defn{dependency-correct}: every column is a partial order (i.e.\ is acyclic/antisymmetric).
\end{enumerate}
\end{definition}
